---
title: "Survival Analysis"
author: "Jaeho Hoya Lee"
output:
  html_document:
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---
Setup data 
```{r setup, include=FALSE}
##set up
getwd()
setwd("/Users/jaehohoyalee/Downloads")
heartdata<-read.csv('heart_failure_clinical_records_dataset.csv',encoding="UTF-8")
str(heartdata)
table(heartdata$DEATH_EVENT) #0->203,1->96

with(heartdata,hist(age))
```
Loading Library
```{r}
library('survival') ##package for survival analysis
library('ggplot2')  ##package for survival visualization
library('survminer')##package for survival visualization
```
##Kaplan Meier
####-Create categorical variable from continuous variable
```{r cars}
heartdata$sodiumc <- ifelse(heartdata$serum_sodium >135 & heartdata$serum_sodium<145,
                            "Serum Sodium Normal", "Serum Sodium Abnormal")
heartdata$efraction <-ifelse(heartdata$ejection_fraction<=75 & heartdata$ejection_fraction>=41, 
                             "Ejection Normal", "Ejection Abnormal")
heartdata$sodiumc<- as.factor(heartdata$sodiumc)
heartdata$efraction<- as.factor(heartdata$efraction)
```

```{r}
fit_sd<-survfit(Surv(time,DEATH_EVENT)~sodiumc, data=heartdata)
ggsurvplot(fit_sd,data=heartdata,xlab="Days", ggtheme=theme_minimal())
```
##Final model for risk factors of Kaplan Meier
```{r}
heartdata$riskgp <-ifelse(heartdata$age<65
                          & heartdata$efraction=="Ejection Normal"
                          & heartdata$sodiumc=="Serum Sodium Normal", "Risk Low", "Risk High")
table(heartdata$riskgp)##High: 272, Low: 27
heartdata$riskgp<- as.factor(heartdata$riskgp)
fit<-survfit(Surv(time,DEATH_EVENT)~riskgp, data=heartdata)
ggsurvplot(fit,data=heartdata,xlab="Days", ggtheme=theme_minimal())
```
##Log rank of P-value for risk group
```{r}
survdiff(Surv(time, DEATH_EVENT) ~ riskgp, data=heartdata)
```
####Difference between group seems significant engough, p= 0.009

###Cox proportional Hazard model 
```{r}
cox.risk<-coxph(Surv(time,DEATH_EVENT) ~ riskgp, data=heartdata)
summary(cox.risk)
```
```{r}
Individual_data=with(heartdata,data.frame(riskgp=c("Risk Low", "Risk High")))
fit0<- survfit(cox.risk,newdata=Individual_data)
riskfactor_plot=ggsurvplot(fit0,
                           Individual_data,
                 legend.labs=c("Risk Low","Risk High"),
                 ggtheme = theme_minimal())
riskfactor_plot
```
##Cox proportional Hazard model
```{r}
heart.cox<-coxph(Surv(time,DEATH_EVENT)~age+sex+
                   anaemia+serum_sodium+creatinine_phosphokinase+
                   ejection_fraction+diabetes+high_blood_pressure+
                   smoking+serum_creatinine,data=heartdata)
summary(heart.cox)
```
##Result for hazard ratio visualization of Cox proportional Hazard model
```{r}
ggforest(heart.cox,data=heartdata)
```
```{r}
hf_plot=ggsurvplot(survfit(heart.cox,data=heartdata),
                   palette = 'skyblue',
                   ggtheme=theme_minimal())
hf_plot
```
##Fitting from our cox-model (high blood pressure relation check)
```{r}
high_blood_pressure_data=with(heartdata,data.frame(high_blood_pressure=c(0,1), anaemia=c(0,0), age=rep(mean(age),2), sex=c(0,0),serum_sodium=rep(mean(serum_sodium),2), ejection_fraction=rep(mean(ejection_fraction),2),creatinine_phosphokinase=rep(mean(creatinine_phosphokinase),2),diabetes=c(0,0),smoking=c(0,0),serum_creatinine=rep(mean(serum_creatinine),2)))
##Strata depending on high_blood_pressure (no:0, yes:1)
fit1=survfit(heart.cox,newdata=high_blood_pressure_data)  
Highbloodpressure_plot=ggsurvplot(fit1, high_blood_pressure_data, legend.labs=c("High-Bloodpressure1","High_Bloodpressure2"),ggtheme = theme_minimal())
Highbloodpressure_plot
```